---
title: "Project2"
author: "girldewhirle"
date: "15 July 2015"
output: html_document
---

***Data Set***

***Exploratory Analysis***

```{r}
# load pre-reqs
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)

# Load the ToothGrowth data and perform some basic exploratory data analyses
data("ToothGrowth")
# Provide a basic summary of the data.
## group by supplement and doses
ToothGrowthv2<-group_by(ToothGrowth,supp,dose)
## calculate mean of each sub group
datasummary<-summarise(ToothGrowthv2,mean(len, na.rm = TRUE),sd(len))
## rename columns
colnames(datasummary)[1]<-"Supplement"
colnames(datasummary)[2]<-"Dose"
colnames(datasummary)[3]<-"Mean_Length"
colnames(datasummary)[4]<-"SD_Length"

kable(datasummary,caption="Table 1: Summary of Sample Means and Standard Deviation by Supplement and Dose",align="l")

## plot of dose/supplement
expplot<-ggplot(data = ToothGrowthv2)
expplot + geom_boxplot(aes(x=supp,y=len)) + facet_grid(.~dose)
expplot + geom_jitter(aes(x=dose,y=len,colour=supp)) +
  geom_smooth(aes(x=dose,y=len,colour=supp),method=lm)
```

***Comparision of Tooth Growth by Supplement and Dose***

# Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

To compare the impact of type of supplement on the results a t test is carried out on each of the pairs of populations given the same dose (note this is not a paired t-test).

```{r}
VC0.5<-filter(ToothGrowthv2,supp=="VC" & dose==0.5)
OJ0.5<-filter(ToothGrowthv2,supp=="OJ" & dose==0.5)

VC1.0<-filter(ToothGrowthv2,supp=="VC" & dose==1.0)
OJ1.0<-filter(ToothGrowthv2,supp=="OJ" & dose==1.0)

VC2.0<-filter(ToothGrowthv2,supp=="VC" & dose==2.0)
OJ2.0<-filter(ToothGrowthv2,supp=="OJ" & dose==2.0)

t.test(VC0.5$len,OJ0.5$len)
t.test(VC1.0$len,OJ1.0$len)
t.test(VC2.0$len,OJ2.0$len)

diff0.5<-mean(VC0.5$len)-mean(OJ0.5$len)
diff1.0<-mean(VC1.0$len)-mean(OJ1.0$len)
diff2.0<-mean(VC2.0$len)-mean(OJ2.0$len)
```

```{r,echo=TRUE}
Twoobject<-ggplot(data = VC2.0,aes(x=len))
Twoobject + geom_density() + geom_density(data = OJ2.0,aes(x=len)) +
  geom_vline(xintercept=mean(VC2.0$len,colour="red")) +
  geom_vline(xintercept=mean(OJ2.0$len,colour="green"))
```


***Conclusions***

# State your conclusions and the assumptions needed for your conclusions. 

# Some criteria that you will be evaluated on
# Did you  perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
# Did the student perform some relevant confidence intervals and/or tests?
# Were the results of the tests and/or intervals interpreted in the context of the problem correctly? 
# Did the student describe the assumptions needed for their conclusions?