---
title: "Statistical Inference Analysis of Tooth Growth Data Set"
author: "girldewhirle"
date: "7 June 2015"
output: html_document
---

The data set is taken from a 1946 study - "THE GROWTH OF THE ODONTOBLASTS OF THE INCISOR TOOTH AS A CRITERION OF THE VITAMIN C INTAKE OF THE GUINEA PIG" - the original paper for which can be found [here](http://jn.nutrition.org/content/33/5/491.full.pdf).

It contains 60 observations (a sample of the original data set) of 3 variables:

- length - length of odontoblast cells in microns
- supp - type of supplement given to guinea pig
- dose - amount of supplement given to guinea pig

** Exploratory Analysis **

Prerequists for analysis in RStudio.

```{r,echo=TRUE}
library("plyr")
library("dplyr")
library("tidyr")
library("ggplot2")
```

Plotting graphs of length of tooth split by type of supplement.

```{r,echo=TRUE}
## create plot object
tooth_object<-ggplot(ToothGrowth,aes(x=dose,y=len))
## create plot
tooth_object + geom_point(aes(colour=supp)) + facet_grid(.~supp) + xlab("Dose of Supplement") + ylab("Length of Odontoblast") + ggtitle("Plot of dose of supplement against length, split by supplement")
```

